---
title: 'Published code: Nature Communications dec 2022'
author: "COMMUNITY"
date: "2022-12-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(segmented)
library(readxl)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(splines)
library(effects)
library(Greg)
library(glue)
library(contingencytables)
library(Hmisc)
library(rms)
library(magrittr)
library(Gmisc)
library(olsrr)
library(quantreg)
library(caret)
library(table1)
setHtmlTableTheme("Google")
```

```{r}
```{r build_dataset}
end_of_followup <- as.Date("2022-02-23")
ds <- raw_ds |> 
  select(id = `Fas 1 id`, 
         Age = `Age at Fas 6`,
         Sex,
         sample_date = F6_provdag,
         sero_group, Vaccine3_type,
         study_status = `Nat Com kohort`,
         blood_IgG_WT = `F6_P25_IgG_WT`,
         blood_IgA_WT = `F6_Ser_IgA_Spike (WT)`,
         blood_IgG_BA1 = `F6_P25_IgG_(B1.1.529;BA.1)`,
         serum_ACE2_BA.1.1 = `F6_P25_NEUT_(BA.1.1)`,
         serum_ACE2_BA.2 = `F6_P25_NEUT_(BA.2)`,
         serum_ACE2_WT = `F6_P25_NEUT_A (WT)`,
         NT_WT = WT100_Neutralization, 
         NT_omicron = Omicron100_Neutralization,
         mucosal_IgA_BA1 = `F6_NPH_Korr_IgA_(BA.1)`,
         mucosal_IgA_WT = F6_NPH_Korr_IgA_WT,
         mucosal_IgG_WT = F6_NPH_korr_IgG_WT,
         CT_nadir = Serie_CT_Lowest,
         `symtomatiska dagar`, `Feber 1=ja 2=nej`, `Hosta 1=ja 2=nej`, `Snuva 1=ja 2=nej`, `Lukt/smakbortfall`, `Huvudvärk 1=ja 2=nej`, `Halsont 1=ja 2=nej`,
         Serie_CT_A, Serie_CT_B, Serie_CT_C, Serie_CT_D, Serie_CT_E, Serie_CT_F, Serie_CT_G,
         whole_genome_sequencing = Odling,
         breakthrough_date = `pos första datum i studie`) |> 
  mutate(across(where(is.POSIXt), as.Date),
         across(c(blood_IgG_WT), as.numeric),
         across(starts_with("mucosal") | starts_with("blood") | starts_with("serum"), log2, .names = "{col}_log2"), 
         Infected = factor(study_status,
                           levels = c("Pos", "Pos, 0 symtom", "Inkl"),
                           labels = c("Yes", "Yes", "No")),
         last_observation = case_when(is.na(breakthrough_date) ~ end_of_followup, TRUE ~ breakthrough_date),
         observation_days = as.numeric(last_observation - sample_date),
         observation_weeks = observation_days / 7,
         Sex = factor(Sex, levels = c("Female", "Male")),
         Asymptomatic = factor(`symtomatiska dagar` == "0", levels = c(FALSE, TRUE), labels = c("No", "Yes")),
         prior_infection = factor(sero_group,
                                  levels = c("Sero -", "Possibly -", "Sero +", "Sero -(+)"),
                                  labels = c("No", "No", "Yes", "Yes")), 
         Symtomatic_infection = factor(`study_status` == "Pos", 
                                       levels = c(FALSE, TRUE), 
                                       labels = c("No", "Yes")),
         days_to_viral_clearance = case_when(Serie_CT_A >45 ~ 3, 
                        Serie_CT_B >45 ~ 5, 
                        Serie_CT_C >45 ~ 7,
                        Serie_CT_D >45 ~ 9,
                        Serie_CT_E >45 ~ 11,
                        Serie_CT_F >45 ~ 13,
                        Serie_CT_G >45 ~ 15,
                          TRUE ~ 16),
         NT_WT = as.numeric(NT_WT),
         NT_omicron = as.numeric(NT_omicron)) |>  
  filter(study_status != "Exkl") |> 
  set_column_labels(blood_IgG_WT_log2 = "IgG wt (log2)")



cutoff_values <- list(quartiles_IgG_WT = ds |> 
                        pluck("blood_IgG_WT") |> 
                        quantile(),
                      quartiles_IgG_BA1 = ds |> 
                        pluck("blood_IgG_BA1") |> 
                        quantile(),
                       quartiles_ACE2_WT = ds |> 
                        pluck("serum_ACE2_WT") |> 
                        quantile(),
                      quartiles_ACE2_BA.1.1 = ds |> 
                        pluck("serum_ACE2_BA.1.1") |> 
                        quantile(na.rm = TRUE),
                      quartiles_ACE2_BA.2 = ds |> 
                        pluck("serum_ACE2_BA.2") |> 
                        quantile(na.rm = TRUE),
                      quartiles_mucosal_IgA_WT = ds |> 
                        pluck("mucosal_IgA_WT") |> 
                        quantile(na.rm = TRUE),
                      quartiles_mucosal_IgG_WT = ds |> 
                        pluck("mucosal_IgG_WT") |> 
                        quantile(na.rm = TRUE),
                      quartiles_serum_IgA_WT = ds |> 
                        pluck("blood_IgA_WT") |> 
                        quantile(na.rm = TRUE),
                      quartiles_total_IgA = ds |> 
                        pluck("F6_NPH_Total_IgA") |> 
                        quantile(na.rm = TRUE),
                      quartiles_total_IgG = ds |> 
                        pluck("F6_NPH_Total_IgG") |> 
                        quantile(na.rm = TRUE))
```

```{r prepare regression data set}
reg_data <- ds |> 
  select(Infected, Age, Sex, sero_group, prior_infection, id,
         blood_IgG_WT_log2, blood_IgA_WT_log2, blood_IgG_BA1_log2, serum_ACE2_WT_log2,
         blood_IgG_WT, blood_IgA_WT, blood_IgG_BA1, serum_ACE2_WT, serum_ACE2_BA.1.1, serum_ACE2_BA.2,
         NT_WT, NT_omicron, mucosal_IgA_WT, mucosal_IgG_WT,
         Asymptomatic, CT_nadir, whole_genome_sequencing, Symtomatic_infection,
         `Feber 1=ja 2=nej`, `Hosta 1=ja 2=nej`, `Snuva 1=ja 2=nej`, `Lukt/smakbortfall`, `Huvudvärk 1=ja 2=nej`, `Halsont 1=ja 2=nej`,
         days_to_viral_clearance,
         observation_days,
         observation_weeks) |> 
  mutate(Age_decade = Age / 10,
         upper_quartile_s_IgG_WT = factor(blood_IgG_WT >= cutoff_values$quartiles_IgG_WT["75%"], 
                                     levels = c(FALSE, TRUE), 
                                     labels = c("No", "Yes")),
         upper_quartile_s_IgA_WT = factor(blood_IgA_WT >= cutoff_values$quartiles_serum_IgA_WT["75%"], 
                                     levels = c(FALSE, TRUE), 
                                     labels = c("No", "Yes")),
         upper_quartile_IgG_BA1 = factor(blood_IgG_BA1 >= cutoff_values$quartiles_IgG_BA1["75%"], 
                                     levels = c(FALSE, TRUE), 
                                     labels = c("No", "Yes")),
         upper_quartile_ACE2_WT = factor(serum_ACE2_WT >= cutoff_values$quartiles_ACE2_WT["75%"], 
                                     levels = c(FALSE, TRUE), 
                                     labels = c("No", "Yes")),
         upper_quartile_ACE2_BA.1.1 = factor(serum_ACE2_BA.1.1 >= cutoff_values$quartiles_ACE2_BA.1.1["75%"], 
                                     levels = c(FALSE, TRUE), 
                                     labels = c("No", "Yes")),
         upper_quartile_ACE2_BA.2 = factor(serum_ACE2_BA.2 >= cutoff_values$quartiles_ACE2_BA.2["75%"], 
                                     levels = c(FALSE, TRUE), 
                                     labels = c("No", "Yes")),
        upper_quartile_m_IgA_WT = factor(mucosal_IgA_WT >= cutoff_values$quartiles_mucosal_IgA_WT["75%"], 
                                     levels = c(FALSE, TRUE), 
                                     labels = c("No", "Yes")),
        upper_quartile_m_IgG_WT = factor(mucosal_IgG_WT >= cutoff_values$quartiles_mucosal_IgG_WT["75%"], 
                                     levels = c(FALSE, TRUE), 
                                     labels = c("No", "Yes"))) |> 

```

```{r possion regression}
poisson_regression <- glm(Infected == "Yes" ~ Age_decade + Sex  +
                            upper_quartile_m_IgA_WT + blood_IgG_WT_log2 + prior_infection,
                          offset = log(observation_weeks),
                          data = reg_data,
                          family = poisson)

linear_poisson_regression <- poisson_regression |> 
    update(. ~ . -upper_quartile_IgG_WT + blood_IgG_WT_log2)

nonlinear_poisson_regression <- linear_poisson_regression |> 
  addNonlinearity(variable = "blood_IgG_WT_log2",
                  spline_fn = "ns")


# Nested models: 
poisson_regression_without_SIgG <- update(poisson_regression, .~ . - blood_IgG_WT_log2)

poission_regression_without_prior_infection <- update(poisson_regression, .~ . - prior_infection)
  
poission_regression_withouth_prior_infection_mIgA <- update(poisson_regression, .~ . - prior_infection - upper_quartile_m_IgA_WT)

poission_regression_without_mIgA <- update(poisson_regression, .~ . - upper_quartile_m_IgA_WT)


#prior infected only
poisson_regression_hybrid <- update(poisson_regression, .~ . - prior_infection, 
                                    subset = prior_infection == "Yes")

#naive only
poisson_regression_naive <- update(poisson_regression, .~ . - prior_infection, 
                                    subset = prior_infection == "No")

#risk of symptomatic infection
poisson_regression_symptomatic <- glm(Symtomatic_infection == "Yes" ~ Age_decade + Sex + prior_infection + blood_IgG_BA1_log2,
                          offset = log(observation_weeks),
                          data = reg_data,
                          family = poisson)


printCrudeAndAdjustedModel(poisson_regression_naive,
                           add_references = TRUE,
                           reference_zero_effect = 1)


```

```{r goodness of fit of nested models}

lrtest(poisson_regression, poission_regression_without_prior_infection)
# p= 0.3295

lrtest(poission_regression_without_mIgA, poission_regression_withouth_prior_infection_mIgA)
# p=0.02

lrtest(poission_regression_without_prior_infection, poission_regression_withouth_prior_infection_mIgA)
# p= 0.0008

lrtest(poisson_regression, poisson_regression_without_SIgG)
# p=0.03 
```

```{r regression nadir Ct}
Ct_model <- glm(CT_nadir ~ Age + Sex + prior_infection + blood_IgG_WT_log2 + Asymptomatic, data = ds)

printCrudeAndAdjustedModel(Ct_model)


# Check how WGS result affects Ct_nadir
Ct_model_wgs <- lm(CT_nadir ~ Age + Sex + prior_infection + blood_IgG_WT_log2 + whole_genome_sequencing + Asymptomatic, data = reg_data)

#check for normality

plot(Ct_model)
hist(Ct_model$residuals, main = "Residual Histogram")

```
WGS is lifted out from the Ct-regression since it cannot be performed on those >30, excluding the ones with high nadir levels. 


```{r regression time to viral clearance}

time_to_clearance_model <- lm(days_to_viral_clearance ~ Age + Sex + prior_infection + blood_IgG_WT_log2 + Asymptomatic, data = reg_data) 

time_to_clearance_wgs <- lm(days_to_viral_clearance ~ Age + Sex + prior_infection + blood_IgG_WT_log2 + whole_genome_sequencing, data = ds)

printCrudeAndAdjustedModel(time_to_clearance_model, add_references = TRUE)

```

